---
title: "Mueller_DS202_Lab2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#All defaults
choco <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/data/choco.csv") 
str(choco)
```


## 1. How are the ratings distributed? Create both a histogram and a boxplot and describe them. Don’t forget to mention outliers, if there are any.

```{r}
#All defaults
plot(choco$Rating)
boxplot(choco$Rating)
```

The ratings are distributed to each REF. The REF is the reference number of the chocolate. There are approximately four outliers that have very low ratings, but other than that the ratings range from 2.5 to 4.0. The average of the ratings is around 3.25.

## 2. How does the number of ratings depend on the location of the company? Draw a bar chart of the number of reports.

```{r}
#All defaults
reports <- table(choco$Location)
str(reports)
best5 <-head(sort(reports,decreasing=TRUE),n=5)
barplot(best5)
```

The five locations with the most ratings are U.S.A., France, Canada, U.K., and Italy. The United States has around four times the amount of ratings than each of the other four countries. This may have to do with population size, since the U.S.A. significantly has more people than the other four countries.

## 3. Do ratings depend on the cacao percentage of a chocolate bar? Create both numerical and graphical summaries, and explain. [Hint: Think about why the scatterplot is not a very good choice for visualization here.]

```{r}
#All defaults
cor(choco$Cocoa_Percent, choco$Rating)
boxplot(choco$Cocoa_Percent ~ choco$Rating)
```
It appears that the percentage of the cacao powder in a chocolate bar has a negative correlation with the rating. According to the box plot, the chocolate bars that have a large average percentage of cacao powder in them typically have lower ratings than the ones that have a lower amount of cacao powder in them. All the higher rated chocolate bars have an average of 70 cacao powder in them.

## 4. How do different bean origins compare in terms of the flavor of chocolates? Are beans from some places better than the other? Create both numerical and graphical summaries, and explain.

```{r}
#All defaults
origin <- table(choco$Bean_Origin)
str(origin)
origin5 <- head(sort(origin,decreasing=TRUE),n=5)
names(origin5)
barplot(origin5)
```

```{r}
#All defaults
venezuela <- choco[choco$Bean_Origin == 'Venezuela',]
peru <- choco[choco$Bean_Origin == 'Peru',]
ecuador <- choco[choco$Bean_Origin == 'Ecuador',]
dominican <- choco[choco$Bean_Origin == 'Dominican Republic',]
madagascar <- choco[choco$Bean_Origin == 'Madagascar',]
venrating <- venezuela$Rating
perurating <- peru$Rating
ecrating <- ecuador$Rating
domrating <- dominican$Rating
madrating <- madagascar$Rating
venave <- mean(venrating)
peruave <- mean(perurating)
ecave <- mean(ecrating)
domave <- mean(domrating)
madave <- mean(madrating)
averating <- c(venave, peruave, ecave, domave, madave)
names(averating) <- c('Venezuela','Peru','Ecuador','Dominican Republic','Madagascar')
barplot(averating)
```

```{r}
#All defaults
print(averating)
```

The five bean origins that had the most entries were Venezuela, Peru, Ecuador, Dominican Republic and Madagascar. As you can see by the numerical and graphical summaries, the average rating of the chocolates whose beans originated from these countries are very similar. They all had a rating of around 3.2. Out of these five origins, Madagascar had the highest average rating (3.26) while Ecuador had the lowest average rating (3.15). From this analysis, I don't think that the origin of the beans has a lot to do with the quality of the chocolate.

## 5. Brainstorm two analyses you would like to perform (with your teammates). Describe in plain language what comparisons are you making. Create suitable numerical and graphical summaries for your analyses. Describe the results.

Our first analysis will be determining if the companies have any affect on rating. First, we will extract the 5 companies that have the most entries in the data set. Then, we will extract the rating data of these companies from the data set. After that, we will create a bar graph and print the average ratings. 

```{r}
#All defaults
company <- table(choco$Company)
str(company)
company5 <- head(sort(company,decreasing=TRUE),n=5)
names(company5)
barplot(company5)
```
```{r}
#All defaults
soma <- choco[choco$Company == 'Soma',]
arete <- choco[choco$Company == 'Arete',]
fresco <- choco[choco$Company == 'Fresco',]
bonnat <- choco[choco$Company == 'Bonnat',]
pralus <- choco[choco$Company == 'Pralus',]
somarating <- soma$Rating
areterating <- arete$Rating
frescorating <- fresco$Rating
bonnatrating <- bonnat$Rating
pralusrating <- pralus$Rating
somaave <- mean(somarating)
areteave <- mean(areterating)
frescoave <- mean(frescorating)
bonnatave <- mean(bonnatrating)
pralusave <- mean(pralusrating)
averating <- c(somaave, areteave, frescoave, bonnatave, pralusave)
names(averating) <- c('Soma','Arete','Fresco','Bonnat','Pralus')
barplot(averating)
```
```{r}
#All defaults
print(averating)
```

According to my data analysis, the five companies with the most entries are Soma, Arete, Fresco, Bonnat, and Pralus. These companies had more varied ratings than the bean origins did in problem 4. The highest rated company was Soma with an average rating of 3.59 and the lowest rating was Pralus with a 3.17. From this analysis, I think that the company does have an impact on rating, but only slightly.

For our second analysis, we are going to determine whether or not the date that the chocolate was made impacts the rating. We are going to take the five dates that have the most entries. Then, we will pull their rating information from the data set and calculate the average rating for each of these dates. We will make a graph and print the data.

```{r}
#All defaults
date <- table(choco$Date)
str(date)
date5 <- head(sort(date,decreasing=TRUE),n=5)
names(date5)
barplot(date5)
```
```{r}
#All defaults
fifteen <- choco[choco$Date == '2015',]
fourteen <- choco[choco$Date == '2014',]
eighteen <- choco[choco$Date == '2018',]
sixteen <- choco[choco$Date == '2016',]
nineteen <- choco[choco$Date == '2019',]
rating2015 <- fifteen$Rating
rating2014 <- fourteen$Rating
rating2018 <- eighteen$Rating
rating2016 <- sixteen$Rating
rating2019 <- nineteen$Rating
ave2015 <- mean(rating2015)
ave2014 <- mean(rating2014)
ave2018 <- mean(rating2018)
ave2016 <- mean(rating2016)
ave2019 <- mean(rating2019)
dateaverating <- c(ave2015, ave2014, ave2018, ave2016, ave2019)
names(dateaverating) <- c('2015','2014','2018','2016','2019')
barplot(dateaverating)
```
```{r}
#All defaults
print(dateaverating)
```

The five dates with the most entries in the data set are 2015, 2014, 2018, 2016, and 2019. The average ratings for all five of these years are fairly similar, they are all around 3.2. The highest rated date is 2015 with an average rating of 3.24, and the lowest rated year was 2019 with an average rating of 3.13. From this analysis, I don't think that the year a chocolate was made has any impact on its flavor.