---
title: "team"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#All defaults
choco <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/data/choco.csv") 
str(choco)
```

## Madalyn Analysis

Our first analysis will be determining if the companies have any affect on rating. First, we will extract the 5 companies that have the most entries in the data set. Then, we will extract the rating data of these companies from the data set. After that, we will create a bar graph and print the average ratings. 

```{r}
#All defaults
company <- table(choco$Company)
str(company)
company5 <- head(sort(company,decreasing=TRUE),n=5)
names(company5)
barplot(company5)
```

```{r}
#All defaults
soma <- choco[choco$Company == 'Soma',]
arete <- choco[choco$Company == 'Arete',]
fresco <- choco[choco$Company == 'Fresco',]
bonnat <- choco[choco$Company == 'Bonnat',]
pralus <- choco[choco$Company == 'Pralus',]
somarating <- soma$Rating
areterating <- arete$Rating
frescorating <- fresco$Rating
bonnatrating <- bonnat$Rating
pralusrating <- pralus$Rating
somaave <- mean(somarating)
areteave <- mean(areterating)
frescoave <- mean(frescorating)
bonnatave <- mean(bonnatrating)
pralusave <- mean(pralusrating)
averating <- c(somaave, areteave, frescoave, bonnatave, pralusave)
names(averating) <- c('Soma','Arete','Fresco','Bonnat','Pralus')
barplot(averating)
```

```{r}
#All defaults
print(averating)
```

According to my data analysis, the five companies with the most entries are Soma, Arete, Fresco, Bonnat, and Pralus. These companies had more varied ratings than the bean origins did in problem 4. The highest rated company was Soma with an average rating of 3.59 and the lowest rating was Pralus with a 3.17. From this analysis, I think that the company does have an impact on rating, but only slightly.

## Sydney Analysis

Our second analysis will be to determine if the years/date of the chocolate have any affect on rating. Our plan is to first extract the top 5 characteristics, from the Characteristic1 variable, that have the most amount of entries in the data set. Then we will extract the rating data of these 5 characteristics from the data set. After that, we will create a bar graph and print the average ratings for each characteristic. 

```{r}
#All defaults
dates <- table(choco$Date)
str(dates)
dates5 <- head(sort(dates,decreasing=TRUE),n=5)
names(dates5)
barplot(dates5)
```

```{r}
five <- choco[choco$Date == '2015',]
four <- choco[choco$Date == '2014',]
eight <- choco[choco$Date == '2018',]
six <- choco[choco$Date == '2016',]
nine <- choco[choco$Date == '2019',]
fiverating <- five$Rating
fourrating <- four$Rating
eightrating <- eight$Rating
sixrating <- six$Rating
ninerating <- nine$Rating
fiveave <- mean(fiverating)
fourave <- mean(fourrating)
eightave <- mean(eightrating)
sixave <- mean(sixrating)
nineave <- mean(ninerating)
averating2 <- c(fiveave, fourave, eightave, sixave, nineave)
names(averating2) <- c('2015','2014','2018','2016','2019')
barplot(averating2)
averating2
```

The top 5 years shown in the graphical bar plot include 2015, 2014, 2018, 2016 and 2019. We can also see from the bar plot that 2015 was the year with the most entries from the data  set. Continuing on with the analysis, we can see from the second bar plot and numerical list that the average ratings for the top 5 years are very similar meaning that the years/date do not have an affect on the chocolate ratings.
